Tutorial: SQL-on-Hadoop Systems

### 2015-04-02 paper.tex
File formats

There is a wide variety of solutions, system architectures, and capabilities in this space, with varying degree of SQL support and capabilities. The purpose of this tutorial is to provide an overview of these options, categorize and compare them to gain insights to open research problems.

@@

Even though the driver that enables all the system presented in this tutorial to work in a single cluster environment is the shared file system, Hadoop File System (HDFS), there is actually a variety of file format that are being used to store data. 
By far the most popular file format is of course plain text.
But there are other options, most of them being variants of PAX \cite{AilamakiDHS01}. For example, Avro, sequence, RC, ORC and DRWF, as well as Parquet.
We present them and perform a comparison between them in terms of usability, data size (compression) and speed of processing

>>>>

There is a wide variety of solutions, system architectures, and capabilities in this space, with varying degree of SQL support and capabilities. The purpose of this tutorial is to provide an overview of these options, categorize and compare them to gain insights into open research problems.

@@

Even though the driver that enables all the system presented in this tutorial to work in a single cluster environment is the shared file system, Hadoop File System (HDFS), there is actually a variety of file format that are being used to store data. 
By far the most popular file format is of course {\em plain text}.
But there are other options.

{\em Avro} uses JSON for defining data types and protocols, and compacts data in a binary format. 
{\em SequenceFiles} are flat files of key/value pairs.
While popular there are some column-oriented, more accurately PAX-like \cite{AilamakiDHS01}, file formats: {\em RC}, {\em ORC} and {\em DWRF} \cite{dwrf}, as well as {\em Parquet}.
We present them and perform a comparison between them in terms of usability, data size (compression) and speed of processing
